### 2015-05-03 internal_review_response.Rnw

Page 3: Would bet a lot of readers won't have any idea what this means. Since the swmpr object is central to the package, maybe a section describing its structure and a line or two on S3 might be a good idea.

{\it Agreed.  This material was already in the text but not as a separate section.  I added a section describing the object structure, including the following text for clarity: `The \texttt{swmpr} object class was created for use with the organizing and analyzing functions.  This object-oriented approach is standard for R (i.e., the S3 object  system, [16]), such that specific methods for a generic function are developed for the object class.}

Page 4: Why only 100? web service limit?

{\it I've modified all functions that have limits on the number of records that can be downloaded, so this is no longer relevant.}

@@

Page 4: Same as above. If there are service limitations should outline those. Just saw this explained below!

{\it Same as above.}

Page 4: This is going to be a tough one to keep relevant. What's the reason for including it?

{\it Agreed, the description of this function was removed from the manuscript but it is retained in the package.  I think it's useful to include but shouldn't be advertised as a primary means of importing data.}

@@

Page 4: Couldn't you build this functionality into the function itself? Perhaps next version... No need to include prior to publishing this.

{\it The most recent version now includes this functionality.}

@@

Page 5: Might be more informative to show the results for each request... dplyr style?

Page 5: If you built in the loop to get around the limit would this not be necessary? This seems like a bandaid that will be less and less relevant over time.

{\it Yes, I've removed this function from the manuscript as noted above.}

### 2015-05-03 swmpr_manu.Rnw
Revisions for changes in CDMO import functions

Three approaches can be used to import \ac{SWMP} data into R, either through direct download or by importing local data (\Cref{tab:retrieve}). First, functions from the package can be used to import the data directly from the online server using \ac{CDMO} web services. To do so, the IP address for the computer making the request must be registered with \ac{CDMO}.\footnote{Follow instructions here: \url{http://cdmo.baruch.sc.edu/webservices.cfm}}  The \texttt{site\_codes} or \texttt{site\_codes\_ind} functions can be used to view the available metadata after a computer is registered with \ac{CDMO}.

% table for retrieval functions
<<results = 'asis', echo = FALSE>>=
funcs <- c('all\\_params', 'all\\_params\\_dtrng', 'import\\_local', 'import\\_remote', 'single\\_param', 'site\\_codes', 'site\\_codes\\_ind')
funcs <- paste0('\\texttt{', funcs, '}')

descrips <- c(
  'Retrieve up to 100 records\\textsuperscript{\\textit{a}} starting with the most recent at a given station, all parameters.  Wrapper to \\texttt{exportAllParamsXMLNew} function on web services.',
  'Retrieve up to 1000 records of all parameters within a given date range for a station.  Optional argument for a single parameter. Wrapper to \\texttt{exportAllParamsDateRangeXMLNew}.', 
  'Import files from a local path.  The files must be in a specific format, such as those returned from the \\ac{CDMO} using the zip downloads option.',
  'Import site data from an independent server. These files have been downloaded from \\ac{CDMO} up to December 2014, processed using SWMPr functions, and uploaded to an Amazon server for quicker import into R.',
  'Retrieve up to 100 records for a single parameter starting with the most recent at a given station.  Wrapper to \\texttt{exportSingleParamXMLNew} function on web services.', 
  'Metadata for all stations, wrapper to \\texttt{exportStationCodesXMLNew} function on web services.',
  'Metadata for all stations at a single site, wrapper  to \\texttt{NERRFilterStationCodesXMLNew} function on web services.'
  )

to_tab <- data.frame(Functions = funcs, Description = descrips, stringsAsFactors = F)
foot <- '\\textsuperscript{\\textit{a}} \\footnotesize Server limits prevent retrieving more records in a single request.'

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = 'Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?all\\_params} at the command line).',
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  insert.bottom = foot,
  label = 'tab:retrieve', 
  table.env = FALSE
  )
@

@@

Due to rate limitations on the \ac{CDMO} server, the retrieval functions return a limited number of records.  The functions are more useful for evaluating short time periods, although these functions could be used iteratively (i.e., with \texttt{for} loops) to obtain longer time series.  Data retrieval functions to access the \ac{CDMO} include \texttt{all\_params}, \texttt{all\_params\_dtrng}, and \texttt{single\_param}: \texttt{all\_params} returns the most recent records of all parameters at a station, \texttt{all\_params\_dtrng} returns all records within a date range for all parameters or a single parameter, and \texttt{single\_param} is identical to \texttt{all\_params} except that a single parameter is requested.

@@

A second approach for importing \ac{SWMP} data into R uses the \texttt{import\_remote} function.  This method imports site data from an independent server where \ac{SWMP} data have been downloaded previously and uploaded as \texttt{swmpr} objects in RData binary format.  This approach is advantageous because the files can be imported very quickly and an IP address does not need to be registered.  The files include almost all available \ac{SWMP} data for each site up to December 2014.  However, the files will not be regularly updated and \ac{CDMO} should always be used for the most recent and accurate data.  The files can be accessed by copying the URL to a web browser with the site name and file extension (e.g., \href{https://s3.amazonaws.com/swmpalldata/acebbnut.RData}{https://s3.amazonaws.com/swmpalldata/acebbnut.RData}, see the \nameref{supp_info}) or by using the function as follows.

<<eval = F>>=
# import a file from an independent server
import_remote('acebbwq')
@

A third approach for data retrieval is to use the \texttt{import\_local} function to import data into R that are locally available after downloading from \ac{CDMO}.  This approach is most appropriate for large, specific data requests. The \texttt{import\_local} function is designed for data from the zip downloads\footnote{\url{http://cdmo.baruch.sc.edu/aqs/zips.cfm}} feature in the advanced query section of the \ac{CDMO}. The function may also work using data from the data export system,\footnote{\url{http://cdmo.baruch.sc.edu/get/export.cfm}} although this feature has not been extensively tested.  The zip downloads feature is an easy way to obtain data from multiple stations in one request.  The downloaded data will be in a compressed folder that includes multiple .csv files by year for a given data type (e.g., apacpwq2002.csv, apacpwq2003.csv, apacpnut2002.csv, etc.).  The \texttt{import\_local} function can be used after the folder is decompressed.

>>>>

Two approaches can be used to import \ac{SWMP} data into R, either through direct download or by importing local data (\Cref{tab:retrieve}). First, functions from the package can be used to import the data directly from the online server using \ac{CDMO} web services. To do so, the IP address for the computer making the request must be registered with \ac{CDMO}.\footnote{Follow instructions here: \url{http://cdmo.baruch.sc.edu/webservices.cfm}}  The \texttt{site\_codes} or \texttt{site\_codes\_ind} functions can be used to view the available metadata after a computer is registered with \ac{CDMO}.

% table for retrieval functions
<<results = 'asis', echo = FALSE>>=
funcs <- c('all\\_params', 'all\\_params\\_dtrng', 'import\\_local', 'single\\_param', 'site\\_codes', 'site\\_codes\\_ind')
funcs <- paste0('\\texttt{', funcs, '}')

descrips <- c(
  'Retrieve records starting with the most recent at a given station, all parameters.  Wrapper to \\texttt{exportAllParamsXMLNew} function on web services.',
  'Retrieve records of all parameters within a given date range for a station.  Optional argument for a single parameter. Wrapper to \\texttt{exportAllParamsDateRangeXMLNew}.', 
  'Import files from a local path.  The files must be in a specific format, such as those returned from the \\ac{CDMO} using the zip downloads option.',
  'Retrieve records for a single parameter starting with the most recent at a given station.  Wrapper to \\texttt{exportSingleParamXMLNew} function on web services.', 
  'Metadata for all stations, wrapper to \\texttt{exportStationCodesXMLNew} function on web services.',
  'Metadata for all stations at a single site, wrapper  to \\texttt{NERRFilterStationCodesXMLNew} function on web services.'
  )

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = 'Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?all\\_params} at the command line).',
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  label = 'tab:retrieve', 
  table.env = FALSE
  )
@

@@

Data retrieval functions to import data directly into R from the \ac{CDMO} include \texttt{all\_params}, \texttt{all\_params\_dtrng}, and \texttt{single\_param}: \texttt{all\_params} returns the most recent records of all parameters at a station, \texttt{all\_params\_dtrng} returns all records within a date range for all parameters or a single parameter, and \texttt{single\_param} is identical to \texttt{all\_params} except that a single parameter is requested. Due to rate limitations on the \ac{CDMO} server, the retrieval functions return a limited number of records with each request.  However, the SWMPr functions use the native CDMO web services iteratively (i.e., within a loop) to obtain the desired time series.  Download time can be excessive for longer time series.

@@

The second approach for data retrieval is to use the \texttt{import\_local} function to import data into R that are locally available after downloading from \ac{CDMO}.  This approach is most appropriate for large, specific data requests. The \texttt{import\_local} function is designed for data from the zip downloads\footnote{\url{http://cdmo.baruch.sc.edu/aqs/zips.cfm}} feature in the advanced query section of the \ac{CDMO}. The zip downloads feature can be used to obtain data from multiple stations in one request.  The downloaded data will be in a compressed folder that includes multiple .csv files by year for a given data type (e.g., apacpwq2002.csv, apacpwq2003.csv, apacpnut2002.csv, etc.).  The \texttt{import\_local} function can be used after the folder is decompressed.

### 2015-05-03 swmpr_manu.tex
Revisions for changes in CDMO import functions

Three approaches can be used to import \ac{SWMP} data into R, either through direct download or by importing local data (\Cref{tab:retrieve}). First, functions from the package can be used to import the data directly from the online server using \ac{CDMO} web services. To do so, the IP address for the computer making the request must be registered with \ac{CDMO}.\footnote{Follow instructions here: \url{http://cdmo.baruch.sc.edu/webservices.cfm}}  The \texttt{site\_codes} or \texttt{site\_codes\_ind} functions can be used to view the available metadata after a computer is registered with \ac{CDMO}.

% table for retrieval functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?all\\_params} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), insert.bottom = foot,     label = "tab:retrieve", table.env = FALSE)%
\begin{table}[!tbp]
\caption{Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?all\_params} at the command line).\label{tab:retrieve}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{all\_params}&Retrieve up to 100 records\textsuperscript{\textit{a}} starting with the most recent at a given station, all parameters.  Wrapper to \texttt{exportAllParamsXMLNew} function on web services.\tabularnewline
\texttt{all\_params\_dtrng}&Retrieve up to 1000 records of all parameters within a given date range for a station.  Optional argument for a single parameter. Wrapper to \texttt{exportAllParamsDateRangeXMLNew}.\tabularnewline
\texttt{import\_local}&Import files from a local path.  The files must be in a specific format, such as those returned from the \ac{CDMO} using the zip downloads option.\tabularnewline
\texttt{import\_remote}&Import site data from an independent server. These files have been downloaded from \ac{CDMO} up to December 2014, processed using SWMPr functions, and uploaded to an Amazon server for quicker import into R.\tabularnewline
\texttt{single\_param}&Retrieve up to 100 records for a single parameter starting with the most recent at a given station.  Wrapper to \texttt{exportSingleParamXMLNew} function on web services.\tabularnewline
\texttt{site\_codes}&Metadata for all stations, wrapper to \texttt{exportStationCodesXMLNew} function on web services.\tabularnewline
\texttt{site\_codes\_ind}&Metadata for all stations at a single site, wrapper  to \texttt{NERRFilterStationCodesXMLNew} function on web services.\tabularnewline
\hline
\end{tabular}\end{center}

\textsuperscript{\textit{a}} \footnotesize Server limits prevent retrieving more records in a single request.\end{table}

@@

A second approach for importing \ac{SWMP} data into R uses the \texttt{import\_remote} function.  This method imports site data from an independent server where \ac{SWMP} data have been downloaded previously and uploaded as \texttt{swmpr} objects in RData binary format.  This approach is advantageous because the files can be imported very quickly and an IP address does not need to be registered.  The files include almost all available \ac{SWMP} data for each site up to December 2014.  However, the files will not be regularly updated and \ac{CDMO} should always be used for the most recent and accurate data.  The files can be accessed by copying the URL to a web browser with the site name and file extension (e.g., \href{https://s3.amazonaws.com/swmpalldata/acebbnut.RData}{https://s3.amazonaws.com/swmpalldata/acebbnut.RData}, see the \nameref{supp_info}) or by using the function as follows.

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import a file from an independent server
> import_remote('acebbwq')
\end{verbatim}
\end{kframe}
\end{knitrout}

A third approach for data retrieval is to use the \texttt{import\_local} function to import data into R that are locally available after downloading from \ac{CDMO}.  This approach is most appropriate for large, specific data requests. The \texttt{import\_local} function is designed for data from the zip downloads\footnote{\url{http://cdmo.baruch.sc.edu/aqs/zips.cfm}} feature in the advanced query section of the \ac{CDMO}. The function may also work using data from the data export system,\footnote{\url{http://cdmo.baruch.sc.edu/get/export.cfm}} although this feature has not been extensively tested.  The zip downloads feature is an easy way to obtain data from multiple stations in one request.  The downloaded data will be in a compressed folder that includes multiple .csv files by year for a given data type (e.g., apacpwq2002.csv, apacpwq2003.csv, apacpnut2002.csv, etc.).  The \texttt{import\_local} function can be used after the folder is decompressed.

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import data
> data(apaebmet)
> dat <- apaebmet
> 
> # select two parameters from dat
> subset(dat, select = c('rh', 'bp'))
> 
> # subset records greater than or equal to a date
> subset(dat, subset = '2013-01-01 0:00', operator = '>=')
> 
> # subset records within a date range
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'))
> 
> # subset records within a date range, select two parameters
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'),
>   select = c('atemp', 'totsorad'))
> 
> # remove rows/columns that do not contain data
> subset(dat, rem_rows = T, rem_cols = T)
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

Two approaches can be used to import \ac{SWMP} data into R, either through direct download or by importing local data (\Cref{tab:retrieve}). First, functions from the package can be used to import the data directly from the online server using \ac{CDMO} web services. To do so, the IP address for the computer making the request must be registered with \ac{CDMO}.\footnote{Follow instructions here: \url{http://cdmo.baruch.sc.edu/webservices.cfm}}  The \texttt{site\_codes} or \texttt{site\_codes\_ind} functions can be used to view the available metadata after a computer is registered with \ac{CDMO}.

% table for retrieval functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \\texttt{?all\\_params} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:retrieve",     table.env = FALSE)%
\begin{table}[!tbp]
\caption{Retrieval functions available from the SWMPr package. Full documentation for each function is in the help file (e.g., execute \texttt{?all\_params} at the command line).\label{tab:retrieve}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{all\_params}&Retrieve records starting with the most recent at a given station, all parameters.  Wrapper to \texttt{exportAllParamsXMLNew} function on web services.\tabularnewline
\texttt{all\_params\_dtrng}&Retrieve records of all parameters within a given date range for a station.  Optional argument for a single parameter. Wrapper to \texttt{exportAllParamsDateRangeXMLNew}.\tabularnewline
\texttt{import\_local}&Import files from a local path.  The files must be in a specific format, such as those returned from the \ac{CDMO} using the zip downloads option.\tabularnewline
\texttt{single\_param}&Retrieve records for a single parameter starting with the most recent at a given station.  Wrapper to \texttt{exportSingleParamXMLNew} function on web services.\tabularnewline
\texttt{site\_codes}&Metadata for all stations, wrapper to \texttt{exportStationCodesXMLNew} function on web services.\tabularnewline
\texttt{site\_codes\_ind}&Metadata for all stations at a single site, wrapper  to \texttt{NERRFilterStationCodesXMLNew} function on web services.\tabularnewline
\hline
\end{tabular}\end{center}

@@

The second approach for data retrieval is to use the \texttt{import\_local} function to import data into R that are locally available after downloading from \ac{CDMO}.  This approach is most appropriate for large, specific data requests. The \texttt{import\_local} function is designed for data from the zip downloads\footnote{\url{http://cdmo.baruch.sc.edu/aqs/zips.cfm}} feature in the advanced query section of the \ac{CDMO}. The function may also work using data from the data export system,\footnote{\url{http://cdmo.baruch.sc.edu/get/export.cfm}} although this feature has not been extensively tested.  The zip downloads feature is an easy way to obtain data from multiple stations in one request.  The downloaded data will be in a compressed folder that includes multiple .csv files by year for a given data type (e.g., apacpwq2002.csv, apacpwq2003.csv, apacpnut2002.csv, etc.).  The \texttt{import\_local} function can be used after the folder is decompressed.

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import data
> data(apaebmet)
> dat <- apaebmet
> 
> # select two parameters from dat
> subset(dat, select = c('rh', 'bp'))
> 
> # subset records greater than or equal to a date
> subset(dat, subset = '2013-01-01 0:00', operator = '>=')
> 
> # subset records within a date range
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'))
> 
> # subset records within a date range, select two parameters
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'),
+   select = c('atemp', 'totsorad'))
> 
> # remove rows/columns that do not contain data
> subset(dat, rem_rows = T, rem_cols = T)
\end{verbatim}
\end{kframe}
\end{knitrout}
